%% This is file `elsarticle-template-1-num.tex',
%% %% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%

%\documentclass[]{article} %\documentclass[final,3p,12pt]{elsarticle}
%\documentclass[final,12pt,times]{elsarticle}

%% Use the option review to obtain double line spacing
\documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{lipsum}
\usepackage{natbib}

\usepackage{tikz}
\usetikzlibrary{shapes, arrows}

\tikzstyle{notary} = [rectangle, draw, text centered, text width=4em, fill=blue!20, minimum height=2em]
\tikzstyle{user}   = [rectangle, draw, text centered, text width=4em, fill=green!20, minimum height=2em]
\tikzstyle{server} = [rectangle, draw, text centered, text width=4em, fill=red!20, minimum height=2em]
\tikzstyle{single}   = [draw, -latex']
\tikzstyle{double}   = [draw, latex-latex']

% http://tex.stackexchange.com/questions/35712/modify-footer-used-by-elsarticle-cls
\makeatletter
\def\ps@pprintTitle{
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{\centerline{\thepage}}
    \let\@evenfoot\@oddfoot}
\makeatother

\journal{University of Guelph; CIS*4110}

\begin{document}

\begin{frontmatter}

\title{Improving the Perspectives Project by creating Notary aware web servers}

\author[doug]{Douglas Anderson}
\author[eric]{Eric Boyd}
\author[james]{James Kelly}
\address[doug]{dander01@uoguelph.ca}
\address[eric]{boyde@uoguelph.ca}
\address[james]{kellyj@uoguelph.ca}


\begin{abstract}

Today's web is secured with SSL and TLS, which utilizes RSA to create secure
connections between servers and users. To ensure that the users are
communicating with the web server they intend and not someone masquerading as
the server, Certificate Authorities (CAs) are considered to be secure trusted
parties that can confirm the identity of the server. While better than a
completely insecure protocol, CAs can be compromised with Man-in-the-Middle
(MitM) attacks and leave users vulnerable. In an effort to fix this weaknesses
in the CA, The Perspectives Project uses a constellation of notary servers
spread out across the Internet to monitor the SSL certificate signatures over
time. Users install a Firefox extension that compares the signature of the
certificate given to them from the server to the certificates observed by the
notaries. Our augmentation to the Perspectives system allows web servers
currently being monitored by a number of notaries to send a message to each of
these notaries when the server changes their SSL certificate. This mitigates
the lag between the certificate change and observation and therefore protects
the users more efficiently.

\end{abstract}

\begin{keyword}
% keywords here, in the form: keyword \sep keyword
Security \sep
SSL \sep
Certificate Authorities \sep
Perspectives
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{intro}

In the modern world the Internet holds a very important role in commerce and
social aspects of life and both of these pursuits require the ability for two
or more parties to communicate securely. To facilitate this secure
communication, the Internet has resorted to using the SSL (Secure Socket Layer)
and it successor TLS (Transport Security Layer) to protect the data being
exchanged.  These protocols utilize public-private key pairs to facilitate RSA
encryption.  While this protocol has been extremely successful it must rely on
a Certificate Authority (CA) to confirm the identity of the owner of public
keys. While it is possible to create `self-signed' certificates, browsers will
warn users if a site uses one of these certificate since it is impossible to
verify who created the certificate. 

Certificate Authorities are points in a network that issues and manages
security credentials and public keys for message encryption. CAs have been
compromised in the past resulting in successful impersonations of several
popular, high profile websites.  In 2011 the central authority Comodo Group,
Inc. was hacked. The attack was easily executed because of Comodo's extremely
weak password system that was easily broken with a word list. The hacker made
off with the SSL certificates for various sites including Gmail, Yahoo Mail,
Hotmail.  \citep{comodohack} This allowed the hacker to perform
Man-in-the-Middle (MitM) attacks on these sites by impersonation the CAs for
the various websites. It allowed the hacker to trick the site user's computers
into believing the fake SSL certificates were genuine and the hacker was then
able to read the site user's emails.  This illustrates the vulnerability in the
SSL protocol that central authorities create.

This vulnerability is what the Perspectives Project sets out to fix. Their
solution to removing central authorities is to create trusted services known as
notaries. These notaries work in tandem with a browser extension that allows
the user of the extension to select which of the notaries they would like to
trust. The system works by checking the SSL certificate signature the user is
given from the server against the SSL certificate signatures that the user's
trusted notaries have in their databases. If there is a discrepancy between the
certificate signature that the server has returned and the signatures that the
notaries have on record, the SSL connection is aborted.

\begin{figure}[h]
\begin{tikzpicture}[node distance = 4cm, auto]
    \node[server](server){Server};
    \node[notary, right of=server](note){Notary};
    \node[user, below of=server](user){User};
    \path[double](note) -- node {Notary Queries \& Replies}(user);
    \path[double](user) -- node {Normal SSL connection}(server);
    \path[single](note) -- node {Probing}(server);
\end{tikzpicture}
\caption{Normal Performance of the perspectives system.}
\end{figure}

\section{Implementation}
\label{implementation}

For our project we attempted to build upon the Perspective Project by allowing
servers to send messages to notaries requesting that each notary acknowledge
the change of the SSL certificate. This fixes the problem in the Perspectives
Project that can occur when a server changes its SSL certificate and a user
queries the notary for that server's SSL certificate. Because the notary has
not yet updated its record of the SSL certificate the notary reports to the
user that there is a mismatch which may lead the user to abort their attempt to
create an SSL connection.

Notary web-crawlers make themselves known to web servers, by setting a custom
HTTP user-agent string, similar to how Google and other search engine's
web-crawlers function. If web servers take note of the notaries checking their
certificates, by checking for this user-agent string, they know how recent
their SSL certificate entry is within the notary network system. In our
implementation the notaries user agent string was:

\begin{verbatim}
    Perspectives-Notary/4.4
\end{verbatim}

\begin{figure}[h]
\begin{tikzpicture}[node distance = 4cm, auto]
    \node[server](server){Server};
    \node[notary, right of=server](note){Notary};
    \node[user, below of=server](user){User};
    \path[double](note) -- node {Notary Queries \& Replies}(user);
    \path[double](user) -- node {Normal SSL connection}(server);
    \path[single](note) -- node {Probing}(server);
    \path[->](server) edge[very thick, bend left=10] node{Probe Request}(note);
\end{tikzpicture}
\caption{Our modified perspective notary allows for the server to request that
    its SSL certificate be probed and updated in the Notaries Database.}
\end{figure}

When a server changes it's SSL certificate, it sends an HTTP request to all of
the notaries it is aware of. In our implementation, we simply have the webserver
load a webpage on the notary servers, that triggers that notary to rescan it's 
certificate and update it's records appropriately. The request is in the following 
form:

\begin{verbatim}
    http://<Notary URL>/?host=<Server URL>&update=certificate
\end{verbatim}



\section{Results}
\label{results}

Our testing was confined to a single web server and two notary servers. Our
results were positive, updating the notary servers within seconds on receiving
a probe request. It is hard to directly compare this to characteristics of the
original notaries because each notary updates itself on a schedule. If a notary
probes its known sites once every day as recommended in the Perspectives Project's
documentation, and the server changes its SSL certificate shortly after it is 
probed, the notary could be reporting the wrong certificate for up to a full day.
When comparing several hours to several seconds, it becomes clear how creating 
notary aware servers is a great benefit for the integrity of the system.

\section{Conclusion}
\label{conclusion}

Our augmentation of the Perspectives system helps protect users by creating a
method to keep Notaries more up to date, which allows the system to do a better
job of validating signatures by reducing the false negative rate. This tendency
for Notaries to incorrectly refute a site's SSL certificate signature because
it is slow to update it's records is an issue with the current Perspectives
implementation. By adding the mechanisms to create notary aware servers, we
have improved the integrity of the system.

\section{Further Work}
\label{further work}

\subsection{Add preventive measures against DOS}

Since it is relatively computationally expensive for the notary to probe a
server for a SSL certificate signature it is possible to use these requests to
make a particular notary unresponsive to user queries. This could allow an
attacker to perform a man in the middle attack by rendering all of a users
notaries unresponsive before beginning their attack. While it may be possible
to prevent this attack on the part of the user by not continuing with the SSL
connection when none of their trusted notaries respond, This would still be
undesirable since it would prevent the user from making any secure connections.

\begin{enumerate}
    \item {Only probe a server when the probe request comes from the server itself}

        By only probing a server when the request comes from the server itself
        the possibility of external abuse is extremely limited. This method may
        be hard to implement in practise because it is possible for the
        attackers to spoof the origin of probe request to be the IP of the
        server.

    \item {Rate limit probe requests}

        By only allowing a certain number of probe requests in a period of time
        the notary can ensure that it does not compromise it's ability to
        respond to user request. This a common DOS prevention technique because
        of how effective it is, however it is possible that this method would
        become swamped at certain times since many websites change their SSL
        certificate at times of low traffic (e.g. 3AM).
\end{enumerate}

\subsection{Test our implementation more rigorously}

Because of the limited time frame of the project we did not have an opportunity
to test our implementation as rigorously as we would have liked. Being a
computer security project the robustness of the system is of utmost importance.
Some further work could include expanding the system to many Notary aware
servers and many Notaries that respond to probing requests.

Our Implementation would also benefit from testing with various types of
malicious users and malicious servers. This could include clients using user
agent string spoofing in order to have servers notify them when it changes it's
SSL certificate.  While it is unclear whether this has any negative
implications to security it is an avenue that should be further investigated.

\section{References}
\label{references}
\nocite{*}
\bibliographystyle{plainnat}
\bibliography{report}

\end{document}

